The effects of host galaxy properties on merging compact binaries detectable by LIGO

### 2016-08-02 paper/mnras/paper.tex
small changes

\author[R. O'Shaughnessy et al.]{R. O'Shaughnessy,$^1$\thanks{E-mail: oshaughn@mail.rit.edu}
J. M. Bellovary,$^2$ 
C. R. Christensen,$^3$ 
S. Shen,$^4$ 
F. Governato,$^5$ 
A. Brooks$^6$\\
%
%
%% List of institutions
$^1$Center for Computational Relativity and Gravitation, Rochester Institute of Technology, Rochester, NY 14623, USA  \\
$^2$Department of Physics, Queensborough Community College, Bayside, NY 11364, USA\\
$^3$Department of Physics, Grinnell College, Grinnell, IA, 50112, USA\\
$^4$Institute of Astronomy, University of Cambridge, Cambridge CB3 OHA, UK\\
$^5$Department of Astronomy, University of Washington, Seattle, WA 98195, USA\\
$^6$Department of Physics and Astronomy, Rutgers University, Piscataway, New Jersey, 08854, USA
 }

@@

  Due to the differences in merger histories, the star formation histories and metallicity
  evolution of h258 and h277 also differ at early times.  Figure
  \ref{fig:TwoGalaxies} shows the star formation history (left panel) and metallicity evolution (right panel) of h277 (black) and h258 (red).  The star formation histories are quite different at early times, where h277 has larger bursts of star formation between 2-4 Gyr, but h258 has a large burst at $\sim 6$ Gyr during a major merger.   The right panel shows the mean metallicity of
  recently formed stars vs time (thick solid lines), where we define
  recent as within the past 50 Myr.  The shaded regions correspond to
  one standard deviation of the mean, while the thick dashed lines
  represent the top and bottom 90\%.  We see that the early evolution
  of the metal properties of these galaxies does differ - between 0.5
  and 6 Gyr, h277 hosts a \emph{modestly}  more metal-rich population than
  h258.   For most astrophysical processes, the small metallicity difference illustrated here will have no impact on
  present-day observables.  However,  the population of binary black holes  depends sensitively on all low metallicity star formation
  over cosmic time; as we show below, these small differences can have an observable impact at present.

@@

To explore plausible binary detection scenarios, we adopt a parametrized formalism for binary evolution and event
detection in an individual galaxy, motivated by
the detailed studies of  \abbrvPSgrb{} (\cite{PSgrbs-popsyn}) and \abbrvPSellipticals{} (\cite{PSellipticals});  similar approaches have been used by
\cite{2016arXiv160508783L} and others.

@@

%\editremark{recheck numbers: we are tweaking things, and Zsun is seemingly inconsistent in the two approaches }
 Our two Milky Way-like galaxies have extremely similar star formation histories and metallicity evolution, particularly at late
 times.  
However, at early times, the \ExcitingGalaxy{} galaxy forms stars for $\simeq 2\unit{Gyr}$ at a lower characteristic metallicity
($Z\simeq 0.8 \times10^{-3}$) compared
to the \BoringGalaxy{} galaxy ($Z\simeq 2\times 10^{-3}$).  In this regime, the formation efficiency $\lambda$ and sensitive volume $V$
can depend sensitively on mass; for example, for binary black holes, the ratio of  $(\lambda V)_{\BoringGalaxy{}}/(\lambda
V)_{\ExcitingGalaxy{}} \simeq 3$.\footnote{Adopting a more extreme exponent for the metallicity dependence ($a=2$) only
  changes this ratio by of order $2$.}   However, in this same epoch, the star formation rate in the \ExcitingGalaxy{} (low-metallicity)
galaxy is smaller, by a factor of roughly 2.
% 1e-3:  V = 475, lambda= 20   
% 2e-3:  V= 318, lambda = 10
% so the ratio should be a factor of about 7x higher due to this region.
Therefore, because only a fraction of order $10\%$  of all star formation occurs in this epoch, by this order of magnitude
argument, the overall number of
present-day coalescing binary black holes in our two galaxies is expected to differ by of order ten percent.
% 0.9 + 0.1*7/2
%
Using the concrete phenomenological calculations described above, we in fact find
$(r_{D}/M)_{\BoringGalaxy{}}/(r_{D}/M)_{\ExcitingGalaxy{}}\simeq 0.9$.

>>>>

\author[R. O'Shaughnessy et al.]{R. O'Shaughnessy,$^1$\thanks{E-mail: oshaughn@mail.rit.edu}
J. M. Bellovary,$^2$ 
A. Brooks$^3$
C. R. Christensen,$^4$ 
S. Shen,$^5$ 
F. Governato,$^6$ \\
%
%
%% List of institutions
$^1$Center for Computational Relativity and Gravitation, Rochester Institute of Technology, Rochester, NY 14623, USA  \\
$^2$Department of Physics, Queensborough Community College, Bayside, NY 11364, USA\\
$^3$Department of Physics and Astronomy, Rutgers University, Piscataway, New Jersey, 08854, USA\\
$^4$Department of Physics, Grinnell College, Grinnell, IA, 50112, USA\\
$^5$Institute of Astronomy, University of Cambridge, Cambridge CB3 OHA, UK\\
$^6$Department of Astronomy, University of Washington, Seattle, WA 98195, USA\\
 }

@@

  Due to the differences in merger histories, the star formation histories and metallicity
  evolution of h258 and h277 also differ at early times.  Figure
  \ref{fig:TwoGalaxies} shows the star formation history (left panel) and metallicity evolution (right panel) of h277 (black) and h258 (red).\footnote{We choose not to show galaxies h239 and h285 in our figures, as they fall in between the values bracketed by h258 and h277 and add confusion to the plots.}  The star formation histories are quite different at early times, where h277 has larger bursts of star formation between 2-4 Gyr, but h258 has a large burst at $\sim 6$ Gyr during a major merger.   The right panel shows the mean metallicity of
  recently formed stars vs time (thick solid lines), where we define
  recent as within the past 50 Myr.  The shaded regions correspond to
  one standard deviation of the mean, while the thick dashed lines
  represent the top and bottom 90\%.  We see that the early evolution
  of the metal properties of these galaxies does differ - between 0.5
  and 6 Gyr, h277 hosts a \emph{modestly}  more metal-rich population than
  h258.   For most astrophysical processes, the small metallicity difference illustrated here will have no impact on
  present-day observables.  However,  the population of binary black holes  depends sensitively on all low metallicity star formation
  over cosmic time.%; as we show below, these small differences can have an observable impact at present.

@@

To explore plausible binary detection scenarios, we adopt a parametrized formalism for binary evolution and event
detection in an individual galaxy, motivated by
the detailed studies of  \cite[][(hereafter \abbrvPSgrb{})]{PSgrbs-popsyn} and  \citet[][(hereafter \abbrvPSellipticals{})]{PSellipticals};  similar approaches have been used by
\cite{2016arXiv160508783L} and others.

@@

%\editremark{recheck numbers: we are tweaking things, and Zsun is seemingly inconsistent in the two approaches }
 Our four Milky Way-like galaxies have extremely similar star formation histories and metallicity evolution, particularly at late
 times.  
However, at early times, the \ExcitingGalaxy{} galaxy forms stars for $\simeq 2\unit{Gyr}$ at a lower characteristic metallicity
($Z\simeq 0.8 \times10^{-3}$) compared
to the \BoringGalaxy{} galaxy ($Z\simeq 2\times 10^{-3}$).  In this regime, the formation efficiency $\lambda$ and sensitive volume $V$
can depend sensitively on mass; for example, for binary black holes, the ratio of  $(\lambda V)_{\BoringGalaxy{}}/(\lambda
V)_{\ExcitingGalaxy{}} \simeq 3$.\footnote{Adopting a more extreme exponent for the metallicity dependence ($a=2$) only
  changes this ratio by of order $2$.}   However, in this same epoch, the star formation rate in the \ExcitingGalaxy{} (low-metallicity)
galaxy is smaller, by a factor of roughly 2.
% 1e-3:  V = 475, lambda= 20   
% 2e-3:  V= 318, lambda = 10
% so the ratio should be a factor of about 7x higher due to this region.
Therefore, because only a fraction of order $10\%$  of all star formation occurs in this epoch, by this order of magnitude
argument, the overall number of
present-day coalescing binary black holes in our two galaxies is expected to differ by of order ten percent.
% 0.9 + 0.1*7/2
%
Using the concrete phenomenological calculations described above, we in fact find
$(r_{D}/M)_{\BoringGalaxy{}}/(r_{D}/M)_{\ExcitingGalaxy{}}\simeq 0.9$.    \jillianremark{\bf RICHARD this is 0.1 in my calcuations}
